\documentclass[a4paper,11pt,reqno]{amsart}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{booktabs}
\usepackage{newtxtext,newtxmath}
\usepackage{microtype}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{enumitem}
\usepackage{needspace}
\usepackage{etoolbox}
\usepackage{xurl}
\usepackage[hidelinks,pdfencoding=auto,psdextra]{hyperref}
\usepackage{bookmark}

\setcounter{secnumdepth}{1}
\setcounter{tocdepth}{1}
\setlength{\parindent}{1.5em}
\setlength{\parskip}{0pt}
\setlength{\emergencystretch}{3em}
\setlist[itemize]{leftmargin=2em,itemsep=0.12em,topsep=0.35em}
\setlist[enumerate]{leftmargin=2.2em,itemsep=0.12em,topsep=0.35em}
\allowdisplaybreaks[2]
\raggedbottom
\Urlmuskip=0mu plus 1mu
\frenchspacing
\tolerance=1800
\pretolerance=500
\ifdefined\pdfgentounicode
  \pdfgentounicode=1
\fi

\theoremstyle{plain}
\newtheorem*{theoremA}{Theorem A}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\BeforeBeginEnvironment{theoremA}{\Needspace{8\baselineskip}}
\BeforeBeginEnvironment{theorem}{\Needspace{6\baselineskip}}
\BeforeBeginEnvironment{proposition}{\Needspace{6\baselineskip}}
\BeforeBeginEnvironment{lemma}{\Needspace{6\baselineskip}}
\BeforeBeginEnvironment{corollary}{\Needspace{5\baselineskip}}
\BeforeBeginEnvironment{conjecture}{\Needspace{6\baselineskip}}
\BeforeBeginEnvironment{definition}{\Needspace{5\baselineskip}}
\BeforeBeginEnvironment{remark}{\Needspace{4\baselineskip}}
\BeforeBeginEnvironment{proof}{\Needspace{4\baselineskip}}

\newcommand{\F}{\mathbb F_5}
\newcommand{\one}{\mathbf 1}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\PhiScore}{\Phi}

\hypersetup{
  pdftitle={Exact Counterexamples and Spectral Mechanisms for WOW-284},
  pdfauthor={Samuil Petkov},
  pdfsubject={Exact 38-, 39-, 40-, 42-, and 50-vertex counterexamples to the minimum-dual-degree and least-distance-eigenvalue conjecture WOW-284},
  pdfkeywords={WOW-284, Moore graph, Hoffman-Singleton graph, (6,5)-cage, distance spectrum, dual degree, girth, counterexample, spectral graph theory},
  pdfcreator={LaTeX with amsart},
  pdfproducer={pdfTeX},
  pdfdisplaydoctitle=true
}

\title[Exact counterexamples to WOW-284]
{Exact Counterexamples and Spectral Mechanisms for WOW-284}
\author{Samuil Petkov}
\address{Department of Physics, \'Ecole normale sup\'erieure,
Universit\'e PSL, Paris, France}
\email{samuil.petkov@phys.ens.psl.eu}
\date{}
\subjclass[2020]{Primary 05C50; Secondary 05C12, 05E30}
\keywords{distance spectrum, dual degree, girth, Moore graph,
Hoffman--Singleton graph, \((6,5)\)-cage, WOW-284, strongly regular graph}

\begin{document}

\begin{abstract}
WOW-284 asserts that the minimum dual degree of every connected graph of
girth at least five is at most the negative of its least distance
eigenvalue.  We give explicit counterexamples of orders
\(38,39,40,42\), and \(50\), all obtained from the
Hoffman--Singleton graph.  The 40-vertex example is 6-regular and has
distance spectrum
\[
  \{75^{(1)},3^{(5)},0^{(16)},(-5)^{(18)}\},
\]
while deleting the endpoints of an edge gives a 38-vertex graph with
\[
  \delta^*=\frac{17}{3},
  \qquad
  \lambda_{\min}(D)=-3-\sqrt7.
\]
More generally, for every connected \(k\)-regular graph of girth at
least five and diameter three,
\[
 D=3J+(k-3)I-2A-A^2,
\]
which yields an exact adjacency-spectral criterion for violating
WOW-284.  Moore second subconstituents provide a further parameterized
construction.  All finite certificates use exact arithmetic, and the
explicit examples have independently checked Lean 4.31 certificates.
\end{abstract}

\maketitle

\section*{Introduction}

Let \(G\) be a finite simple connected graph on \(n\ge2\) vertices.  Its
distance matrix is \(D(G)=(d_G(u,v))_{u,v\in V(G)}\), where \(d_G\) is graph
distance.  This matrix is real symmetric; order its real eigenvalues as
\[
  \partial_1(G)\ge \partial_2(G)\ge\cdots\ge\partial_n(G).
\]
For a vertex \(v\), its \emph{dual degree} is the average degree of its
neighbors,
\[
  d^*(v)=\frac{1}{d(v)}\sum_{u\in N(v)}d(u),
  \qquad
  \delta^*(G)=\min_{v\in V(G)}d^*(v).
\]
Write \(g(G)\) for the shortest-cycle length, with \(g(G)=\infty\) for an
acyclic graph; spectral superscripts indicate algebraic multiplicity.

Aouchiche and Hansen record the following Graffiti conjecture in their survey
of distance spectra~\cite[Conjecture~7.16, p.~370]{AouchicheHansen2014}; they
attribute it to Fajtlowicz's 1998 \emph{Written on the Wall}
report~\cite{Fajtlowicz1998}.

\begin{conjecture}[WOW-284]\label{conj:wow284}
If \(G\) is connected, has \(n\ge3\) vertices, and has girth
\(g(G)\ge5\), then
\[
  \delta^*(G)\le -\partial_n(G).
\]
\end{conjecture}

For such a graph, write
\[
  \Phi(G):=\delta^*(G)+\partial_{|V(G)|}(G).
\]
Thus \(G\) is a strict counterexample precisely when \(\Phi(G)>0\).

In that 2014 survey, the authors reported that the conjecture remained open
and noted equality for the Petersen
graph~\cite[Conjecture~7.16, p.~370]{AouchicheHansen2014}.  This equality is the
degree-3 boundary case of a general Moore-graph calculation.  The degree-7
Moore graph constructed by Hoffman and
Singleton~\cite{HoffmanSingleton1960} lies beyond that boundary and gives a strict
violation.

There is important prior spectral work.  Howlader and Panigrahi determine a
distance polynomial for minimal \((k,5)\)-cages and explicitly list the
distance spectra of the Petersen, Hoffman--Singleton, and hypothetical
degree-57 Moore
graphs~\cite[Theorems~2.3 and~2.5(1)]{HowladerPanigrahi2022}; the existence of the
last graph remains open~\cite{SmithMontemanni2026}.  Their \(k=7\) calculation,
together with regularity and the girth condition, already supplies every
ingredient needed to disprove Conjecture~\ref{conj:wow284}.  The present note
records the sharp degree criterion, makes the WOW-284 connection explicit,
and supplies a self-contained coordinate certificate.  It also extracts the
40-vertex \((6,5)\)-cage from the same coordinates and gives its distance
spectrum directly.  No claim is made that the distance spectra are new, that
this observation has priority over every unpublished observation, or that the
38-vertex example has minimum possible order among all counterexamples.

\phantomsection\label{thm:moore-wow}
\begin{theoremA}
Let \(G\) be a Moore graph of diameter two and degree \(k\ge2\).  Then
\[
 |V(G)|=k^2+1,\qquad g(G)=5,\qquad \delta^*(G)=k,
 \qquad
 \partial_{k^2+1}(G)=-\frac{3+\sqrt{4k-3}}2.
\]
Consequently \(G\) satisfies the inequality in WOW-284 if and only if
\(k\le3\), with equality exactly when \(k=3\).
\end{theoremA}

\section{The Moore-graph mechanism}\label{sec:moore}

A \emph{Moore graph of diameter two and degree \(k\)} is a finite simple
connected \(k\)-regular graph of diameter two attaining the Moore bound
\(|V(G)|\le 1+k+k(k-1)=k^2+1\).  We include the short derivation needed for
\hyperref[thm:moore-wow]{Theorem~A}; no classification of Moore graphs is used.

Fix a vertex \(v\).  There are exactly \(k(k-1)\) nonbacktracking walks
\(vuw\) of length two.  There are also exactly \(k(k-1)\) vertices outside
\(\{v\}\cup N(v)\), and diameter two implies that each such vertex is the
endpoint of at least one of these walks.  Walks with distinct endpoints are
distinct, so these required walks already exhaust all \(k(k-1)\)
nonbacktracking walks.  Consequently every such walk ends outside
\(\{v\}\cup N(v)\), and every outside vertex is reached exactly once.
Applying this at every vertex shows that adjacent vertices have no common
neighbor and nonadjacent vertices have exactly one.  Thus there are no
triangles or 4-cycles.  To exhibit a 5-cycle, take an edge \(uv\), choose
\(x\in N(u)\setminus\{v\}\) and \(y\in N(v)\setminus\{u\}\), and let \(z\)
be the unique common neighbor of the necessarily nonadjacent vertices
\(x,y\).  The absence of triangles and 4-cycles makes these five vertices
distinct, so \(uxzyvu\) is a 5-cycle.  Hence \(g(G)=5\).

Let \(A\), \(I\), and \(J\) denote the adjacency, identity, and all-ones
matrices of order \(n=k^2+1\).  The common-neighbor counts give, entry by
entry,
\begin{equation}\label{eq:general-a-square}
 A^2=(k-1)I-A+J.
\end{equation}
Indeed, the diagonal entries are \(k\), the entries indexed by edges are
zero, and the remaining off-diagonal entries are one.  Since
\(A\one=k\one\) and the graph is connected, the eigenvalue \(k\) is simple.
For example, for \(x\in\mathbb R^n\),
\[
 x^{\mathsf T}(kI-A)x
   =\sum_{\{u,v\}\in E(G)}(x_u-x_v)^2,
\]
so equality holds only for constant \(x\).  The real symmetric matrix \(A\)
preserves \(\one^\perp\), and~\eqref{eq:general-a-square} restricts there to
\[
 A^2+A-(k-1)I=0.
\]
Writing \(q=\sqrt{4k-3}\), the two nonprincipal adjacency eigenvalues are
\[
 r=\frac{-1+q}{2},\qquad s=\frac{-1-q}{2}.
\]
If \(m_r,m_s\) are the respective multiplicities, then
\[
  m_r+m_s=k^2,\qquad k+m_rr+m_ss=0.
\]
Solving gives
\begin{equation}\label{eq:general-multiplicities}
 m_r=\frac12\left(k^2+\frac{k(k-2)}q\right),\qquad
 m_s=\frac12\left(k^2-\frac{k(k-2)}q\right).
\end{equation}

Every nonedge joins vertices at distance two, so
\begin{equation}\label{eq:general-distance}
 D=2J-2I-A.
\end{equation}
It follows that the full distance spectrum is
\begin{equation}\label{eq:general-distance-spectrum}
 \Spec(D)=
 \left\{(2k^2-k)^{(1)},
 \left(\frac{q-3}{2}\right)^{(m_s)},
 \left(-\frac{q+3}{2}\right)^{(m_r)}\right\}.
\end{equation}
In particular, the last displayed eigenvalue is the least one.  Regularity
gives \(\delta^*(G)=k\).  Hence the conjectured inequality is
\[
 k\le\frac{3+\sqrt{4k-3}}2.
\]
For \(k\ge2\), comparison after moving 3 to the left is legitimate, and
\[
 (2k-3)^2-(4k-3)=4(k-1)(k-3).
\]
Thus the inequality is strict for \(k=2\), is an equality for \(k=3\), and
fails precisely for \(k>3\).  This proves
\hyperref[thm:moore-wow]{Theorem~A}.

Our smallest explicit example is the 38-vertex graph
\(H=R-\{P_{1,0},P_{1,1}\}\), for which
\[
  \delta^*(H)=\frac{17}{3},
  \qquad
  \partial_{38}(H)=-3-\sqrt7,
  \qquad
  \Phi(H)=\frac83-\sqrt7>0.
\]

\section{Coordinate construction}

All subscripts in this section lie in the field
\(\F=\mathbb Z/5\mathbb Z\).  Introduce vertices
\[
 V(G)=\{P_{i,j}:i,j\in\F\}\mathbin{\dot\cup}
      \{Q_{k,\ell}:k,\ell\in\F\}.
\]
Thus \(|V(G)|=50\).  The edges are the following unordered pairs:
\begin{align*}
 E_P&=\bigl\{\{P_{i,j},P_{i,j+1}\}:i,j\in\F\bigr\},\\
 E_Q&=\bigl\{\{Q_{k,\ell},Q_{k,\ell+2}\}:k,\ell\in\F\bigr\},\\
 E_X&=\bigl\{\{P_{i,j},Q_{k,ik+j}\}:i,j,k\in\F\bigr\}.
\end{align*}
Set \(E(G)=E_P\cup E_Q\cup E_X\).  For a fixed numerical labeling, with
representatives in \(\{0,1,2,3,4\}\), use
\[
 P_{i,j}\longleftrightarrow 5i+j,
 \qquad
 Q_{k,\ell}\longleftrightarrow 25+5k+\ell.
\]

After reindexing the \(Q\)-layers by \(k\mapsto-k\), these are exactly
Hafner's affine-coordinate form of Robertson's pentagon-and-pentagram
construction~\cite[Section~2 and Theorem~2.1]{Hafner2003}.  Thus \(G\) is the
Hoffman--Singleton graph; all properties needed here are proved directly.

The complete neighborhood formulas are
\begin{align}
 N(P_{i,j})
 &=\{P_{i,j-1},P_{i,j+1}\}
   \cup\{Q_{k,ik+j}:k\in\F\},\label{eq:p-neighborhood}\\
 N(Q_{k,\ell})
 &=\{Q_{k,\ell-2},Q_{k,\ell+2}\}
   \cup\{P_{i,\ell-ik}:i\in\F\}.
   \label{eq:q-neighborhood}
\end{align}

\begin{lemma}\label{lem:simple-regular}
The construction defines a simple 7-regular graph with 175 edges.
\end{lemma}

\begin{proof}
Every listed edge has distinct endpoints.  Within each of the five fixed
\(P\)-layers, addition by \(1\) gives a 5-cycle; within each fixed
\(Q\)-layer, addition by \(2\) gives a 5-cycle.  The cross edges have one
endpoint of each type.  The set notation removes any possibility of parallel
edges.

Equations~\eqref{eq:p-neighborhood}--\eqref{eq:q-neighborhood} list seven
distinct neighbors at every vertex.  For a \(P\)-vertex, the two same-layer
neighbors are distinct because \(1\ne-1\) in \(\F\); the five cross
neighbors are distinct because their first \(Q\)-coordinates are distinct;
and the two types cannot overlap.  The \(Q\)-case is identical with steps
\(\pm2\).  There are 25 edges of type \(E_P\), 25 of type \(E_Q\), and 125
cross edges, hence 175 in total.  Equivalently, the handshake identity gives
\(50\cdot7/2=175\).
\end{proof}

\section{The common-neighbor certificate}

\begin{proposition}\label{prop:common-neighbors}
For any two distinct vertices \(x,y\),
\[
 |N(x)\cap N(y)|=
 \begin{cases}
 0,&x\sim y,\\
 1,&x\not\sim y.
 \end{cases}
\]
\end{proposition}

\begin{proof}
First consider two \(P\)-vertices \(P_{i,j}\) and \(P_{i',j'}\).  If
\(i=i'\), their layer is a 5-cycle.  Adjacent pairs have no common
\(P\)-neighbor.  If, for example, \(j'=j+2\), the unique common
\(P\)-neighbor is \(P_{i,j+1}\); the case \(j'=j-2\) is analogous.  They
cannot have a common \(Q\)-neighbor: for fixed \(i\) and \(k\), the equation
\(Q_{k,\ell}\sim P_{i,j}\) determines \(j=\ell-ik\) uniquely.  If
\(i\ne i'\), there is no common \(P\)-neighbor, and a common \(Q\)-neighbor
must satisfy
\[
 ik+j=i'k+j'.
\]
This has the unique solution
\[
 k=\frac{j'-j}{i-i'},\qquad \ell=ik+j,
\]
because \(\F\) is a field.

The argument for two \(Q\)-vertices is dual.  If their first coordinates
agree, they lie in the 5-cycle generated by steps \(\pm2\).  Adjacent pairs
have no common \(Q\)-neighbor.  For a nonadjacent pair, whose second
coordinates differ by \(\pm1\), direct intersection of the two sets of
\(\pm2\)-neighbors gives exactly one common \(Q\)-neighbor; no common
\(P\)-neighbor is possible.  If \(k\ne k'\), a common
\(P_{i,j}\) is determined uniquely by
\[
 \ell=ik+j,\qquad \ell'=ik'+j,
\]
namely
\[
 i=\frac{\ell-\ell'}{k-k'},\qquad j=\ell-ik.
\]

It remains to compare \(P_{i,j}\) with \(Q_{k,\ell}\).  Put
\[
 r=\ell-(ik+j)\in\F.
\]
The pair is adjacent precisely when \(r=0\).  A common \(P\)-vertex must be
\(P_{i,j+s}\) with \(s\in\{\pm1\}\), and it is adjacent to
\(Q_{k,\ell}\) precisely when \(r=s\).  Hence there is exactly one common
\(P\)-neighbor if \(r\in\{\pm1\}\), and none otherwise.  A common
\(Q\)-vertex must be \(Q_{k,\ell+t}\) with \(t\in\{\pm2\}\), and it is
adjacent to \(P_{i,j}\) precisely when \(t=-r\).  Hence there is exactly one
common \(Q\)-neighbor if \(r\in\{\pm2\}\), and none otherwise.  Finally,
\[
 \F=\{0\}\mathbin{\dot\cup}\{\pm1\}
          \mathbin{\dot\cup}\{\pm2\},
\]
which proves every cross case and completes the certificate.
\end{proof}

\begin{corollary}\label{cor:geometry}
The graph is connected, has diameter two, and has girth five.
\end{corollary}

\begin{proof}
Every pair of distinct vertices is adjacent or has a common neighbor, so the
graph is connected and has diameter at most two.  It is not complete, since
it is 7-regular on 50 vertices, so its diameter is exactly two.

A triangle would give an adjacent pair a common neighbor.  In a 4-cycle, two
opposite vertices would have the other two cycle vertices as distinct common
neighbors.  Both possibilities contradict
Proposition~\ref{prop:common-neighbors}.  Thus \(g(G)\ge5\).  On the other
hand, for every fixed \(i\),
\[
 P_{i,0}P_{i,1}P_{i,2}P_{i,3}P_{i,4}P_{i,0}
\]
is a 5-cycle, so \(g(G)=5\).
\end{proof}

\section{The 50-vertex coordinate counterexample}

The graph just constructed is 7-regular on \(50=7^2+1\) vertices and has
diameter two.  It is therefore a Moore graph to which
\hyperref[thm:moore-wow]{Theorem~A} applies.  Substituting \(k=7\)
into~\eqref{eq:general-distance-spectrum} gives
\begin{equation}\label{eq:distance-spectrum}
 \Spec(D)=\{91^{(1)},1^{(21)},(-4)^{(28)}\},
\end{equation}
or equivalently
\[
 \det(tI-D)=(t-91)(t-1)^{21}(t+4)^{28}.
\]
Every vertex and every one of its neighbors has degree 7, so
\[
 d^*(v)=\frac{1}{7}\sum_{u\in N(v)}7=7.
\]
Consequently
\[
 g(G)=5,\qquad \delta^*(G)=7>4=-\partial_{50}(G),
 \qquad \delta^*(G)+\partial_{50}(G)=3.
\]
This proves that the explicitly constructed Hoffman--Singleton graph is a
strict counterexample to WOW-284.

Equivalently, \(D+\delta^*(G)I=D+7I\) has eigenvalues
\(98\), \(8\) with multiplicity 21, and \(3\) with multiplicity 28.  Thus
it is positive definite, which is the matrix form of the strict inequality
\(\delta^*(G)>-\partial_{50}(G)\).

\section{A 40-vertex induced counterexample}\label{sec:forty}

The same coordinates contain a smaller regular counterexample.  Delete
\[
 \mathcal P=\{P_{0,j},Q_{0,j}:j\in\F\}
\]
and let \(R=G-\mathcal P\) be the induced graph on the remaining 40 vertices.
The deleted graph consists of the \(P_{0,\bullet}\) pentagon, the
\(Q_{0,\bullet}\) pentagram, and the matching \(P_{0,j}Q_{0,j}\); hence it
is a Petersen graph.  This construction is the classical 40-vertex
\((6,5)\)-cage of O'Keefe and
Wong~\cite{OKeefeWong1979}.  Its uniqueness was proved by
Wong~\cite{Wong1979}, and its realization as a Petersen deletion from the
Hoffman--Singleton graph is recorded explicitly by Klin, Muzychuk, and
Ziv-Av~\cite[Section~3.6]{KlinMuzychukZivAv2009}.

\begin{theorem}\label{thm:forty}
The graph \(R\) is connected and 6-regular, has girth five and diameter three,
and has distance spectrum
\[
 \Spec(D(R))=\{75^{(1)},3^{(5)},0^{(16)},(-5)^{(18)}\}.
\]
Consequently
\[
 \delta^*(R)=6>5=-\partial_{40}(R),
 \qquad \delta^*(R)+\partial_{40}(R)=1,
\]
so \(R\) is a strict counterexample to WOW-284.
\end{theorem}

\begin{proof}
Every retained vertex has exactly one neighbor in \(\mathcal P\):
\(P_{i,j}\), with \(i\ne0\), is adjacent there only to \(Q_{0,j}\), while
\(Q_{k,\ell}\), with \(k\ne0\), is adjacent there only to \(P_{0,\ell}\).
Thus \(R\) is 6-regular.  It inherits the absence of triangles and 4-cycles
from \(G\), and each retained \(P\)-layer is a 5-cycle, so \(g(R)=5\).

Fix \(v\in V(R)\).  Girth five makes the distance-zero, distance-one, and
distance-two layers from \(v\) have respectively \(1\), \(6\), and
\(6\cdot5=30\) vertices.  Three vertices remain.  Each has no neighbor in
\(L_0(v)\cup L_1(v)\) and has at most two neighbors among the three remaining
vertices; 6-regularity therefore gives it a neighbor in the distance-two
layer.  Hence all three lie at distance three from \(v\).  This proves
connectedness and diameter three.

It remains to compute the spectrum.  Order the vertices of \(G\) with those
of \(R\) first, and write its adjacency matrix in blocks as
\[
 A_G=\begin{pmatrix}B&C\\ C^{\mathsf T}&P\end{pmatrix},
\]
where \(B=A(R)\), \(P=A(G[\mathcal P])\), and \(C\) is \(40\times10\).
Every row of \(C\) has sum one and every column has sum four.  Moreover,
direct inspection of the two 5-cycles and their matching shows that every
nonadjacent pair of vertices in \(\mathcal P\) has exactly one common neighbor
in \(\mathcal P\).  Consequently,
distinct vertices of \(\mathcal P\) have no common retained neighbor: adjacent
pairs have no common neighbor in \(G\), while the unique common neighbor of a
nonadjacent pair already lies in the induced Petersen graph.  Therefore
\begin{equation}\label{eq:c-block}
 C\mathbf1_{10}=\mathbf1_{40},\qquad C^{\mathsf T}C=4I_{10}.
\end{equation}
In particular, \(C\) has rank ten.

Equation~\eqref{eq:general-a-square} with \(k=7\) gives
\(A_G^2=6I-A_G+J\).  Its top-right block is
\[
 BC+CP=-C+J_{40,10}.
\]
Here \(J_{a,b}\) denotes the \(a\times b\) all-ones matrix.  Since
\(J_{40,10}=CJ_{10}\) by~\eqref{eq:c-block},
\begin{equation}\label{eq:bc-intertwining}
 BC=C(J_{10}-I_{10}-P).
\end{equation}
The Petersen adjacency spectrum, obtained from the \(k=3\) case of
Section~\ref{sec:moore}, is
\(\{3^{(1)},1^{(5)},(-2)^{(4)}\}\).  Because \(C\) is injective,
equation~\eqref{eq:bc-intertwining} gives
\[
 \Spec\bigl(B|_{\operatorname{col}C}\bigr)
 =\{6^{(1)},(-2)^{(5)},1^{(4)}\}.
\]

The top-left block of the same identity is
\[
 B^2+CC^{\mathsf T}=6I_{40}-B+J_{40}.
\]
Since \(B\) is symmetric, \(\ker C^{\mathsf T}=(\operatorname{col}C)^\perp\)
is \(B\)-invariant.  If \(x\in\ker C^{\mathsf T}\), then
\(\mathbf1_{40}^{\mathsf T}x=(C\mathbf1_{10})^{\mathsf T}x=0\), so
\((B^2+B-6I)x=0\).  Thus the remaining 30 adjacency eigenvalues are \(2\)
or \(-3\).  If their multiplicities are \(a,b\), then
\(a+b=30\); the displayed ten-dimensional spectrum has trace zero, and
\(\tr B=0\), so \(2a-3b=0\).  Hence \(a=18\), \(b=12\), and
\begin{equation}\label{eq:forty-adjacency-spectrum}
 \Spec(B)=\{6^{(1)},2^{(18)},1^{(4)},(-2)^{(5)},(-3)^{(12)}\}.
\end{equation}

Let \(A_i\) be the distance-\(i\) matrix of \(R\).  Girth five and
6-regularity give \(A_2=B^2-6I\), while diameter three gives
\(J=I+B+A_2+A_3\).  Consequently
\begin{equation}\label{eq:forty-distance-identity}
 D(R)=B+2A_2+3A_3=3J+3I-2B-B^2.
\end{equation}
On \(\mathbf1_{40}\), the eigenvalue is
\(3\cdot40+3-2\cdot6-6^2=75\).  A nonprincipal adjacency eigenvalue
\(\theta\) maps to
\[
 3-2\theta-\theta^2=4-(\theta+1)^2.
\]
Applying this to~\eqref{eq:forty-adjacency-spectrum} gives the claimed
distance spectrum.  Finally, regularity gives \(\delta^*(R)=6\), completing
the proof.
\end{proof}

\section{Smaller induced counterexamples}\label{sec:smaller}

The 40-vertex graph contains strict counterexamples of orders 39 and 38.
All claims in this section are reproduced exactly by the accompanying
programs; floating-point eigenvalues are not used.

\subsection{Deleting the endpoints of an edge}

Delete the adjacent vertices
\[
 a=P_{1,0},\qquad b=P_{1,1}
\]
from \(R\), and call the resulting graph \(H\).  A graph6 string in this fixed
labeling, together with complete adjacency lists and an edge list, is provided
in \texttt{data/graphs/}.

\begin{theorem}\label{thm:thirty-eight}
The graph \(H\) is connected, has 38 vertices, 109 edges, girth five,
diameter three, and degree multiset \(6^{(28)},5^{(10)}\).  Moreover,
\[
 \delta^*(H)=\frac{17}{3},
 \qquad
 \partial_{38}(H)=-3-\sqrt7.
\]
Consequently
\[
 \delta^*(H)+\partial_{38}(H)=\frac83-\sqrt7>0,
\]
so \(H\) is a strict counterexample to WOW-284.
\end{theorem}

\begin{proof}
Induced deletion preserves simplicity and cannot create a triangle or a
4-cycle.  A distinct \(P\)-layer remains a 5-cycle.  Exact breadth-first
search from every vertex gives connectedness and diameter three.

Put
\[
 X=N_R(a)\setminus\{b\},\qquad
 Y=N_R(b)\setminus\{a\}.
\]
The girth condition implies that \(X\) and \(Y\) are disjoint and that no
vertex of \(X\cup Y\) has a remaining neighbor in \(X\cup Y\).  These ten
vertices therefore have degree five and all their remaining neighbors have
degree six; every other vertex has degree six.  A degree-six vertex has at
most one neighbor in \(X\) and at most one in \(Y\), since two in either set
would create a 4-cycle.  If it has \(t\in\{0,1,2\}\) neighbors in
\(X\cup Y\), then
\[
 d^*(v)=\frac{5t+6(6-t)}6=6-\frac{t}{6}.
\]
There are \(10\cdot5=50\) incidences from \(X\cup Y\) to the 28 degree-six
vertices, so some degree-six vertex has \(t=2\).  Hence
\(\delta^*(H)=17/3\).

The exact distance characteristic polynomial is displayed in
Appendix~\ref{app:polynomials}.  It contains \((x^2+6x+2)^2\), whose roots
are \(-3\pm\sqrt7\).  For the square-free part of this polynomial, the exact
Sturm sequence has variation counts
\[
 V(-\infty)=26,\qquad V(-28/5)=25.
\]
Thus the full polynomial has exactly one distinct root below \(-28/5\).
Moreover, \(7>(13/5)^2\), so
\[
 -3-\sqrt7<-3-\frac{13}{5}=-\frac{28}{5}.
\]
Therefore this is the least distance eigenvalue.  Independently, an exact rational
\(LDL^{\mathsf T}\) decomposition of \(3D(H)+17I\) has 38 positive pivots.
Finally \(8/3>\sqrt7\), because \(64>63\), proving strictness.
\end{proof}

\subsection{Orders 39 and 42}

For \(v\in V(R)\), put \(H_v=R-v\).  Deleting \(v\) leaves its six
neighbors with degree five and the other 33 vertices with degree six.
The six degree-five vertices are pairwise nonadjacent, and all five
remaining neighbors of each have degree six; hence their dual degree is
\(6\).  A degree-six vertex has at most one degree-five neighbor, since
two would form a 4-cycle through \(v\).  There are \(6\cdot5=30\)
incidences between the degree-five and degree-six vertices.  Therefore
exactly 30 degree-six vertices have one degree-five neighbor, and
\[
  \delta^*(H_v)=\frac{5+5\cdot6}{6}=\frac{35}{6}.
\]

For the explicit representative
\[
  H_{39}:=R-P_{1,0},
\]
an exact rational \(LDL^{\mathsf T}\) decomposition of
\(6D(H_{39})+35I\) has all 39 pivots strictly positive.  Thus
\(6D(H_{39})+35I\succ0\), so \(H_{39}\) is a strict counterexample.
Proposition~\ref{prop:descendants} below verifies the corresponding
assertion for every choice of \(v\).

For another regular example, let \(X_{42}\) be the graph induced by the
vertices at distance two from \(P_{0,0}\) in the Hoffman--Singleton graph.
The graph \(X_{42}\) is 6-regular and has
\[
 \Spec(A(X_{42}))=\{6^{(1)},2^{(21)},(-1)^{(6)},(-3)^{(14)}\},
\]
\[
 \Spec(D(X_{42}))=\{81^{(1)},4^{(6)},0^{(14)},(-5)^{(21)}\}.
\]
Hence
\(\delta^*(X_{42})=6>5=-\partial_{42}(X_{42})\).
The parameterized construction behind this classical second subconstituent is proved in
Section~\ref{sec:subconstituent}; its adjacency spectrum is also tabulated by
van Dam and Haemers~\cite[Table~3]{vanDamHaemers2003}.

\subsection{Exact labelled descendant families}

\begin{proposition}\label{prop:descendants}
For every vertex \(v\) of \(R\), the graph \(R-v\) is connected, has girth at
least five, has
\(\delta^*=35/6\) and the same exact distance characteristic polynomial
\(P_{39}\) displayed in Appendix~\ref{app:polynomials}.  Every one of these
40 labelled graphs is a strict counterexample.

For every edge \(uv\) of \(R\), the graph \(R-\{u,v\}\) is connected, has
girth at least five, has
\(\delta^*=17/3\), the same exact distance characteristic polynomial
\(P_{38}\), and least distance eigenvalue \(-3-\sqrt7\).  Every one of
these 120 labelled graphs is a strict counterexample.
\end{proposition}

\begin{proof}
The script \texttt{scripts/verify\_descendant\_families.py} enumerates all
40 vertices and all 120 edges.  For every induced graph, breadth-first search
from every vertex checks connectivity and reconstructs the integer distance
matrix.  Induced deletion cannot create a cycle of length below five.  The
script then computes every dual degree as a rational number and the
characteristic polynomial \(\det(xI-D)\) over \(\mathbb Z\).  Exact equality
with \(P_{39}\) or \(P_{38}\) proves the common-polynomial assertions.  The
representative \(H_{39}\) satisfies
\(6D(H_{39})+35I\succ0\), so every root of \(P_{39}\) is greater than
\(-35/6\).  Since every graph \(R-v\) has this same characteristic polynomial,
the same strict bound holds for all 40 labelled deletions.  An exact Sturm
count for \(P_{38}\) identifies its least root as \(-3-\sqrt7\).  No
graph-isomorphism or transitivity assumption is used.
\end{proof}

\begin{remark}
A numerical screen of all \(\binom{40}{3}=9880\) three-vertex deletions found
no graph with \(\Phi>0\); its best value was approximately
\(-0.306979936667\).
This is exploratory evidence only, not an exact elimination of order 37 and
not a minimality proof.
\end{remark}

\section{A diameter-three spectral criterion}\label{sec:diam3}

\begin{theorem}\label{thm:diam3}
Let \(G\) be a connected \(k\)-regular graph on \(n\) vertices, of girth at
least five and diameter three.  Then
\begin{equation}\label{eq:diam3-polynomial}
 D=3J+(k-3)I-2A-A^2.
\end{equation}
The principal distance eigenvalue is \(3n-k^2-k-3\), and every nonprincipal
adjacency eigenvalue \(\theta\) gives the distance eigenvalue
\[
 \mu(\theta)=k-2-(\theta+1)^2.
\]
Consequently \(G\) is a strict counterexample to WOW-284 precisely when
\[
 \max_{\theta\ne k}|\theta+1|<\sqrt{2k-2}.
\]
\end{theorem}

\begin{proof}
The absence of triangles and 4-cycles means that two vertices at distance two
have exactly one common neighbor.  Hence the distance-two matrix is
\(A_2=A^2-kI\).  Since the diameter is three,
\(A_3=J-I-A-A_2\).  Substitution into \(D=A+2A_2+3A_3\) proves
\eqref{eq:diam3-polynomial}.  The asserted eigenvalues follow by applying the
identity to \(\one\) and to \(\one^\perp\).  The matrix \(D\) is nonnegative
and irreducible (indeed, every off-diagonal entry is positive), and its row
sums are constant by the displayed identity.  Perron--Frobenius therefore
makes the principal eigenvalue its largest eigenvalue, so the least distance
eigenvalue occurs among the nonprincipal images.  Finally regularity gives
\(\delta^*=k\), so strict violation is equivalent to
\(k>\max_{\theta\ne k}((\theta+1)^2-k+2)\), which is the displayed
criterion.
\end{proof}

\begin{corollary}[Bipartite obstruction]
A connected \(k\)-regular bipartite graph of diameter three and girth at
least five is not a strict counterexample for \(k\ge3\).
\end{corollary}

\begin{proof}
The nonprincipal adjacency eigenvalue \(-k\) maps to
\[
 \mu(-k)=k-2-(k-1)^2.
\]
Its negative is at least \(k\) for \(k\ge3\), with equality only at \(k=3\).
\end{proof}

For comparison, exact negative controls follow directly from
Theorem~\ref{thm:diam3}.  The Odd graph \(O_4=KG(7,3)\) has adjacency
spectrum
\[
 \{4^{(1)},2^{(14)},(-1)^{(14)},(-3)^{(6)}\},
\]
so \(\delta^*=4\) and \(\partial_n=-7\).  The Heawood graph has adjacency
spectrum
\[
 \{3^{(1)},(\sqrt2)^{(6)},(-\sqrt2)^{(6)},(-3)^{(1)}\},
\]
so \(\delta^*=3\) and \(\partial_n=-2-2\sqrt2\).

\section{Moore second subconstituents}\label{sec:subconstituent}

\Needspace{16\baselineskip}
\begin{theorem}\label{thm:second-sub}
Let \(M\) be a degree-\(K\) Moore graph of diameter two, where \(K\ge3\),
and fix a vertex \(v\).  Let \(X\) be the graph induced by the vertices at
distance two from \(v\).  Then
\[
 |V(X)|=K(K-1),\qquad X\text{ is }(K-1)\text{-regular},
\]
\(X\) has girth at least five and diameter three, and
\[
 \partial_{K(K-1)}(X)=-\frac{5+\sqrt{4K-3}}2.
\]
Consequently
\[
 \delta^*(X)+\partial_{K(K-1)}(X)
 =K-1-\frac{5+\sqrt{4K-3}}2,
\]
which is positive exactly for integers \(K\ge6\).
\end{theorem}

\begin{proof}
Relative to \(\{v\}\sqcup N(v)\sqcup V(X)\), write
\[
 A(M)=
 \begin{pmatrix}
 0&\one^{\mathsf T}&0\\
 \one&0&C\\
 0&C^{\mathsf T}&B
 \end{pmatrix}.
\]
Each vertex of \(X\) has one neighbor in \(N(v)\) and \(K-1\) in \(X\),
so \(B\) is \((K-1)\)-regular.  Distinct rows of \(C\) have disjoint support
and row sum \(K-1\), hence \(CC^{\mathsf T}=(K-1)I\).  Since
\(CC^{\mathsf T}=(K-1)I_K\), the map \(C^{\mathsf T}\) is injective and
\[
  \mathbb R^{V(X)}
  =\operatorname{im}C^{\mathsf T}\mathbin{\oplus}\ker C.
\]
Moreover,
\[
  C^{\mathsf T}\one_K=\one_{V(X)}.
\]
The middle-right block of the Moore identity gives \(CB=J-C\).
Transposing this identity shows that
\[
  BC^{\mathsf T}z=-C^{\mathsf T}z
  \qquad (z\perp\one_K).
\]
Thus \(B\) has eigenvalue \(-1\) on the
\((K-1)\)-dimensional space
\(C^{\mathsf T}(\one_K^\perp)\), while
\(\one_{V(X)}\) has eigenvalue \(K-1\).

Now let \(x\in\ker C\).  Since
\(C^{\mathsf T}\one_K=\one_{V(X)}\),
\[
  \one_{V(X)}^{\mathsf T}x
  =\one_K^{\mathsf T}Cx=0.
\]
Furthermore,
\[
  CBx=(J-C)x=0,
\]
so \(\ker C\) is \(B\)-invariant.  The bottom-right block of the
Moore identity therefore restricts on \(\ker C\) to
\[
 B^2+B-(K-1)I=0,
\]
whose roots are
\[
 r=\frac{-1+\sqrt{4K-3}}2,
 \qquad
 s=\frac{-1-\sqrt{4K-3}}2.
\]
The kernel has dimension \(K(K-2)>0\), so the principal line,
\(C^{\mathsf T}(\one_K^\perp)\), and \(\ker C\) account for all
\[
  1+(K-1)+K(K-2)=K(K-1)
\]
dimensions.  If the multiplicities of \(r,s\)
there are \(m_r,m_s\), then the trace of \(B\) gives
\(m_rr+m_ss=0\), because the principal eigenvalue \(K-1\) cancels the
\((K-1)\) copies of \(-1\).  Since \(r>0>s\), both multiplicities are
positive.

We record the diameter argument explicitly.  Around any \(x\in X\), girth
at least five gives \(1\), \(K-1\), and \((K-1)(K-2)\) distinct vertices in
the first three distance layers of \(X\).  The remaining \(K-2\) vertices are
exactly the other vertices of \(X\) sharing the unique neighbor of \(x\) in
\(N(v)\).  They have no path of length at most two to \(x\), since such a
path would create a triangle or 4-cycle in \(M\).  Every neighbor of any such
remaining vertex \(z\) is nonadjacent to \(x\), since an edge to \(x\) would
complete the 4-cycle through their shared neighbor in \(N(v)\).  If
\(w\in N_X(z)\), the unique ambient common neighbor of \(x\) and \(w\) cannot
lie in \(N(v)\): it would have to be the unique such neighbor of \(x\), and
would then form a triangle with \(w,z\).  It therefore lies in \(X\), giving
a path of length three from \(x\) to \(z\).  Hence all remaining vertices are
at distance three.  This proves connectivity and diameter three.

Apply Theorem~\ref{thm:diam3} with \(k=K-1\).  The eigenvalue \(-1\) maps
to \(K-3\), while \(r\) and \(s\) map respectively to
\[
 -\frac{5+\sqrt{4K-3}}2
 \quad\text{and}\quad
 -\frac{5-\sqrt{4K-3}}2.
\]
Thus the first of these is the least distance eigenvalue.  Finally,
\[
  2\bigl(\delta^*(X)+\partial_{K(K-1)}(X)\bigr)
  =2K-7-\sqrt{4K-3}.
\]
This is nonpositive for \(K=3,4,5\).  For \(K\ge6\), both sides of
\(2K-7>\sqrt{4K-3}\) are positive, and
\[
  (2K-7)^2-(4K-3)
  =4(K^2-8K+13)>0.
\]
Indeed, \(K^2-8K+13=1\) at \(K=6\) and is increasing for \(K\ge6\).
Hence the gap is positive exactly for integers \(K\ge6\).
\end{proof}

For \(K=7\), the two roots on \(\ker C\) are \(2\) and \(-3\).  Their
multiplicities \(m_2,m_{-3}\) satisfy
\[
  m_2+m_{-3}=35,\qquad 2m_2-3m_{-3}=0,
\]
the second equation following from the trace after the principal eigenvalue
\(6\) cancels the six copies of \(-1\).  Hence
\[
  m_2=21,\qquad m_{-3}=14.
\]
This gives the explicit 42-vertex graph above.  A degree-57
Moore graph, if it exists, would give a 56-regular example on 3192 vertices.
Its existence remains open~\cite{SmithMontemanni2026}; the theorem is
therefore a parameterized construction, not an unconditional infinite family.

\section{An equitable-deletion framework}

\begin{proposition}\label{prop:equitable-delete}
Let \(M\) be a degree-\(K\) Moore graph of diameter two.  Partition its
vertices as \(V(M)=U\sqcup S\), write
\[
 A(M)=\begin{pmatrix}B&C\\C^{\mathsf T}&P\end{pmatrix},
\]
and suppose \(M[S]\) is \(r\)-regular while every vertex of \(U\) has
exactly \(t>0\) neighbors in \(S\).  Then \(M[U]\) is
\((K-t)\)-regular.  If \(Py=\eta y\) and \(y\perp\one\), then
\[
 BCy=-(1+\eta)Cy,
 \qquad
 \lVert Cy\rVert^2=(K-1-\eta-\eta^2)\lVert y\rVert^2.
\]
Moreover, on \(\ker C^{\mathsf T}\),
\[
 B^2+B-(K-1)I=0.
\]
Whenever \(M[U]\) has diameter three, these adjacency data feed directly
into Theorem~\ref{thm:diam3}.
\end{proposition}

\begin{proof}
The upper-right and lower-right blocks of the Moore identity are
\[
 BC+CP=-C+J,
 \qquad
 C^{\mathsf T}C+P^2=(K-1)I-P+J.
\]
If \(Py=\eta y\) and \(y\perp\one\), then \(Jy=0\), so the
upper-right block gives
\[
  BCy=-(1+\eta)Cy.
\]
Taking the inner product of the lower-right block with \(y\) gives
\[
  \lVert Cy\rVert^2
  =(K-1-\eta-\eta^2)\lVert y\rVert^2.
\]
Finally, if \(x\in\ker C^{\mathsf T}\), then
\(C\one=t\one\) and \(t>0\) imply \(x\perp\one\).
The upper-left block consequently gives
\[
  (B^2+B-(K-1)I)x=0.
\]
\end{proof}

For the 40-vertex graph, \(U=V(R)\), \(K=7\), \(S\) is the induced Petersen
graph, \(r=3\), and \(t=1\).  If, more generally,
\(S\subsetneq V(M)\) is itself a
degree-\(r\) Moore graph and every outside vertex has exactly one neighbor in
\(S\), cross-edge counting gives
\[
 (r^2+1)(K-r)=K^2-r^2.
\]
Because the outside set is nonempty, \(K>r\), so cancellation yields
\(K=r^2-r+1\).  The coordinate deletions displayed above realize the steps
\(C_5\subset\) Petersen \(\subset\) Hoffman--Singleton, corresponding to
\((r,K)=(2,3)\) and \((3,7)\), but do not yield an infinite chain.

\section{Further exact and exploratory checks}

For \(i\in\mathbb F_5\), put
\(\mathcal L_i=\{P_{i,j},Q_{i,j}:j\in\mathbb F_5\}\), and let \(R_m\) be
the graph obtained from \(G\) by deleting
\(\mathcal L_0,\ldots,\mathcal L_{m-1}\) (with \(R_0=G\)).  These balanced
coordinate-layer deletions give the following exact representatives:
\[
\begin{array}{c@{\qquad}c@{\qquad}c@{\qquad}c@{\qquad}c}
\toprule
m&|V|&\text{degree}&\partial_n&\delta^*+\partial_n\\
\midrule
0&50&7&-4&3\\
1&40&6&-5&1\\
2&30&5&-6&-1\\
3&20&4&-7&-3\\
4&10&3&-3&0\\
\bottomrule
\end{array}
\]
Thus this natural chain stops producing strict counterexamples after the
40-vertex member.  The preceding analytic theorems cover the Moore and
regular diameter-three mechanisms, Moore second subconstituents, and the
equitable deletion framework.  The exact computations check the listed
finite examples, all singleton and edge-endpoint deletions of \(R\), the
balanced layer chain, \(O_4\), and the Heawood graph.

No unconditional infinite family is established.  Fixed-degree,
diameter-three graphs have bounded order by the Moore bound, so an infinite
family within this framework would require unbounded degree.  Covers,
subdivisions, coronas, and general vertex replacements are not asserted to
preserve the strict inequality; the present arguments establish no such
preservation theorem.

\section{Computational certificate}

The accompanying Python programs supply several exact checks with deliberately
separated dependencies.  The original 50-vertex certificate:
\begin{enumerate}
\item constructs the 175 unordered edges and checks all 50 degrees;
\item checks the common-neighbor count for every unordered vertex pair;
\item constructs all 2,500 distance entries by integer breadth-first search;
\item computes the distance characteristic polynomial symbolically; and
\item computes an exact rational \(L\Delta L^{\mathsf T}\) decomposition of
      \(D+7I\) and checks that all 50 pivots are positive.
\end{enumerate}
The resulting characteristic polynomial is
\[
  (t-91)(t-1)^{21}(t+4)^{28}
\]
and all 50 rational pivots are positive.  Writing
\[
  M:=D+7I=L\Delta L^{\mathsf T},
\]
we have, for every nonzero \(y\),
\[
  y^{\mathsf T}My=(L^{\mathsf T}y)^{\mathsf T}
  \Delta(L^{\mathsf T}y)>0,
\]
so this proves positive definiteness of \(M=D+7I\).

A second exact script constructs \(R\) by the stated deletion and independently
checks its 40 degrees, all BFS distances, girth, the block identities
\eqref{eq:c-block}--\eqref{eq:bc-intertwining}, and the characteristic
polynomials
\[
 \det(tI-B)=(t-6)(t-2)^{18}(t-1)^4(t+2)^5(t+3)^{12},
\]
\[
 \det(tI-D(R))=(t-75)(t-3)^5t^{16}(t+5)^{18}.
\]

The extended verifier reconstructs the 42-, 40-, 39-, and 38-vertex graphs
from the field-coordinate rules and checks all structural, dual-degree,
characteristic-polynomial, Sturm, and rational
\(LDL^{\mathsf T}\) assertions used above.  A separate 38-vertex audit starts
from the stored graph6 representation, reconstructs graph distances by
integer breadth-first search through NetworkX~\cite{HagbergSchultSwart2008},
and uses a handwritten \texttt{Fraction}-based decomposition.  This is
independent of the coordinate representation and of SymPy's matrix
factorization, although it is not an independently sourced graph.

Finally, \path{scripts/verify_descendant_families.py} exhausts the 40 labelled
single-vertex and 120 labelled edge-endpoint deletions.  SymPy supplies exact
integer characteristic polynomials \(\det(tI-D)\) and Sturm counts of their
square-free parts~\cite{MeurerEtAl2017}.  The recorded output is
\path{supplement/extended_2026-07-23/logs/descendant_family_output.txt}; the
enclosing \path{SHA256SUMS} records the expected SHA-256 digest of that output
and every other archived certificate, permitting byte-for-byte integrity
checks.
The cross-platform command
\begin{verbatim}
python scripts/verify_extended.py
\end{verbatim}
runs the exact certificates for the explicit non-50 constructions, the
independent graph6 audit, the descendant-family verification, and the
byte-for-byte graph-export checks.  The balanced layer-deletion chain,
the Odd graph, and the Heawood graph are checked by
\begin{verbatim}
python scripts/explore_generalizations.py
\end{verbatim}
which also performs the separately labelled floating-point three-vertex
screen.  No floating-point eigenvalue evidence is used for any theorem.

\section{Formal verification}

The explicit 50-vertex counterexample is fully formalized and verified in
Lean 4.31 with Mathlib 4.31~\cite{deMouraUllrich2021,Mathlib2020}.  The
development verifies the graph, 7-regularity, the exhaustive common-neighbor
certificate, girth five, and the adjacency-square and distance-matrix
identities.  It supplies exact rational matrices \(S,S^{-1}\), proves their
two-sided inverse identities by denominator-cleared integer certificates,
proves \(DS=S\operatorname{diag}(91,(-4)^{(28)},1^{(21)})\), and checks the
multiplicities.  Thus the complete explicit 50-vertex proof, including its
least distance eigenvalue, is kernel-checked.  The scalar \(k\le3\) threshold
is formalized separately.

Lean 4.31 also kernel-checks finite spectral certificates for the explicit
38-, 39-, 40-, and 42-vertex constructions.  For orders 38, 39, and 42, the
public endpoints certify the exact minimum dual-degree data, positive
definiteness of a shifted finite matrix, and the resulting strict inequality
for every nonzero real eigenpair.  For order 40, the endpoint certifies a
two-sided invertible exact diagonalization, its diagonal-entry
multiplicities, and that \(-5\) is an attained minimum diagonal entry, together
with dual degree six and gap one.

The scope of these non-50 results is deliberately finite and spectral: their
public endpoints do not bundle every structural hypothesis, and the semantic
finite matrices are not identified with Mathlib's
\texttt{SimpleGraph.dist} in a single theorem.  They are therefore described
here as kernel-checked finite spectral certificates, rather than as a full
standard-library formalization of every graph-theoretic phrase in the
manuscript.  Section~\ref{sec:moore}, the diameter-three criterion, and the
Moore-subconstituent theorem remain conventional analytic proofs.  The
descendant-family assertions are conventional finite arguments supported by
exact computation.

\section*{Scope, provenance, and disclosures}

The note gives counterexamples of orders 38, 39, 40, 42, and 50.  No claim is
made that 38 is the minimum possible order among WOW-284 counterexamples, and
no exhaustive search over all smaller graphs is part of this note.  Calling
the 40-vertex graph the \((6,5)\)-cage refers to its classical minimum-order
property among 6-regular girth-five graphs, not to minimum order for the WOW
problem.  The distance-spectral input for Moore graphs is credited to
Howlader and Panigrahi; the classical 40-vertex cage and 42-vertex second
subconstituent are separately credited above.  The repository source ledger
records the explicit WOW-284 connection and the limits of the targeted
priority search.

OpenAI ChatGPT assisted with adversarial proof checking, proof exploration,
and Lean formalization.  No AI system is an author.  Samuil Petkov is the
sole named author and assumes full responsibility for the mathematics,
citations, attribution, and conclusions.

The author received no funding for this work and declares no competing
interests.  The manuscript source, exact verification code, machine-readable
certificates, and build instructions are available in the public repository
\href{https://github.com/SamPetkov/wow284}{\texttt{github.com/SamPetkov/wow284}}.
The computational archive corresponding to this manuscript is release
\href{https://github.com/SamPetkov/wow284/releases/tag/v2.0.0-arxiv}
{\texttt{v2.0.0-arxiv}}.

\appendix
\section{Exact distance characteristic polynomials}\label{app:polynomials}

For every labelled single-vertex deletion considered in
Proposition~\ref{prop:descendants},
\begin{align*}
P_{39}(x)={}&x^9(x+5)^{12}(x^2+6x+3)\\
&\cdot(x^3+3x^2-15x-7)^2(x^3+3x^2-15x-3)^2\\
&\cdot(x^4-78x^3+303x^2-70x-450).
\end{align*}
For every labelled edge-endpoint deletion considered there,
\begin{align*}
P_{38}(x)={}&x^4(x-2)(x+5)^8(x^2+6x+2)^2\\
&\cdot(x^3+3x^2-15x-3)\\
&\cdot(x^4+5x^3-7x^2-23x-6)\\
&\cdot(x^5+9x^4+7x^3-77x^2-54x-4)\\
&\cdot(x^9-67x^8-404x^7+1772x^6+7205x^5\\
&\hspace{2.2em}-18489x^4-17018x^3+20288x^2+16824x+1680).
\end{align*}

\begingroup
\footnotesize
\setlength{\bibsep}{0pt}
\renewcommand{\bibsection}{\section*{References}}
\bibliographystyle{unsrtnat}
\bibliography{references}
\endgroup

\end{document}
